---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# wwntests

The goal of the  wwntests package is to provide an array of functional hypothesis tests tests and related visualizations. Currently supported hypothesis tests include weak and strong white noise tests, which are built for general white noise and thus hold under conditional heterscedasticity assumptions, as well as tests of independence and identical distribution.

## Installation

Using the *devtools* package:
```{r}
library(devtools)
devtools::load_all()
```

## CRAN

You can install the released version of wwntests from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("wwntests")
```

## Examples

We provide some basic examples which shows you how to apply the package to functional data. We start with a sequence of Brownian motion curves and a sequence of FAR(1, 0.75)-IID (Functional Auto Regressive Process with IID errors) curves, both of which are available in *wwntests* via the *brown_motion* and *far_1_S* functions. For both functions, *N* determines the number of samples and *J* determines the number of times each individual curve is sampled. The FAR(1,S) process is not a weak white noise (nor a strong white nouse) when S > 0.
```{r}
set.seed(1997)
b <- brown_motion(N = 150, J = 100)
f <- far_1_S(N = 150, J = 100, S = 0.75)
```

We might wish to test whether or not our functional data is a weak-white noise. The 'single-lag' and 'multi-lag' tests test this null hypothesis. For the sake of brevity, we'll look just at the 'single-lag' test, please refer to the vignette for a more in depth look into both tests. 

Note, all tests can be accessed throgh the *fport_test* by passing their string handles (see documentation) to the argument *test*. If you are interested in a particular test, stand-alone functions are also avaiable for each test (see documentation).

To be more specific, the 'single-lag' test at lag h tests the null hypothesis that the lag-h autocovariance function of the data is equal to 0. We apply the single-lag test at lag = 1.
```{r}
# Note, we pass TRUE to the suppress_raw_output to suppress the list output of the function.
fport_test(f_data = b, test = 'single-lag', lag = 1, suppress_raw_output = TRUE)
fport_test(f_data = f, test = 'single-lag', lag = 1, suppress_raw_output = TRUE)
```
Our p-values are as expected, given that a Brownian moton is indeed a weak white noise while the FAR data is not.

Now suppose we want to test whether or not the functional data is independend and identically distributed. The 'spectral' and 'independence' tests test this as a null hypothesis. For brevity, we will give an example using the 'spectral' test, which is based on the spectral density operator of the functional data, and essentially measures the proximity of a functional time series to a white noise - the constant spectral density operator of an uncorrelated series.
```{r}
fport_test(f_data = b, test = 'spectral', kernel = 'Bartlett', bandwidth = 'adaptive', 
           suppress_raw_output = TRUE)
fport_test(f_data = f, test = 'spectral', kernel = 'Bartlett', bandwidth = 'adaptive', 
           suppress_raw_output = TRUE)
```

Again, our p-values are as expected. The 'spectral' and 'independence' test both provide the advantage that they are very quick, much less computationally burdensome than the 'single-lag' and (especially) the 'multi-lag' test. However, the 'spectral' and 'independence' test have the drawback that they are not designed for general white noise and may not hold under functional conditionally heterscedastic assumptions. To see this in action, we will consider an fGARCH(1,1) (Functional Generalized Autoregressive Conditional Heteroscedasticity) process. This process is a weak-white noise, but not a strong white noise, and as a result, we may expect the 'spectral-test' to falsely assert that the fGARCH(1,1) process is independent and identically distributed.
```{r}
fgarch <- fgarch_1_1(N = 150, J = 100)
fport_test(f_data = fgarch, test = 'spectral', kernel = 'Bartlett', bandwidth = 'adaptive',
           suppress_raw_output = TRUE)
```


